\LARGE \bf Imposing a Weight Norm Constraint for Neuro–Adaptive Control

### 2025-03-23 manuscript.tex
final touch before report to Prof.

% \makeatletter
% \let\NAT@parse\undefined
% \makeatother

\input{dding_template/tex_packages/tex_packages_general.tex}
\input{dding_template/tex_macros/tex_macros_general.tex}
% MY PAKEAGES ========================================

\newcommand*{\q}{\mv{q}}
\newcommand*{\dq}{\dot{\q}}
\newcommand*{\ddq}{\ddot{\q}}
\newcommand*{\qd}{\mv{q_d}}
\newcommand*{\dqd}{\dot{\qd}}
\newcommand*{\ddqd}{\ddot{\qd}}

@@

More recently, neuro–adaptive control approaches have been developed using neural networks (NNs) to approximate unknown system dynamics \cite{Farrell:2006aa}.
It uses the approximation capability of NNs with any arbitrary sigmoidal activation function \cite{Cybenko:1989aa}, which can approximate sufficiently smooth functions with minimum approximation error through ideal weights in a compact set.
This property allows the NNs to be synthesized in the controllers to approximate the unknown dynamics.
Various architectures of NNs have been proposed for neuro–adaptive control, such as single-hidden-layer NN (SHLNN) \cite{Esfandiari:2015aa,Gao:2006aa}, radial basis function NN (RBFNN) \cite{Ge:2002aa,Zhou:2023aa} and deep NN (DNN) \cite{Patil:2022aa}.
The literature has shown that NNs can improve the system's performance index and stability by approximating the unknown dynamics.

@@

% One of the common issues in neuro–adaptive control is that the boundedness of the control inputs is not guaranteed.
One of the common issues in using NNs is that the outputs of NNs are not predictable.
This is because the input‒output relationships of the NNs are not interpretable (\ie the NNs are usually called black boxes \cite{Sheu:2020aa,Rudin:2019aa}).
This issue should be addressed for safety because the unpredictable control input may reach the actuator's physical limitations and result in system collapse.

@@

Consider an uncertain Euler‒Lagrange system modeled as
\begin{equation}
    \mm M\ddq + \mm V_m\dq+ \mm G + \mm F 
    =
    \mv\tau 
    ,
    \label{eq. system dynamics 1}
\end{equation}
where $\q\in \mathbb{R}^n$ and $\mv \tau\in\R^n$ denote the generalized coordinate and the control input, respectively; $\mm M :=\mm M(\q)\in\mathbb{R}^{n\times n}$, $\mm V_m:= \mm V_m(\q,\dq)\in\mathbb{R}^{n\times n}$, and $\mm G:=\mm G(\q)\in\mathbb{R}^{n}$ denote the unknown system function matrices; and $\mm F:=\mm F(\dq)\in\mathbb{R}^{n}$ denotes the external force.
Using the user-designed matrices $\widehat{\mm M}>0,\widehat{\mm V_m}$ and $\widehat{\mm G}$, \eqref{eq. system dynamics 1} can be represented as 
\begin{equation}
    \widehat{\mm M}\ddq+\widehat{\mm V_m}\dq+\widehat{\mm G} = \mv \tau + \mv f(\q,\dq,\ddq)
    ,
    \label{eq. system dynamics 2}
\end{equation}
where $\mv f(\q,\dq,\ddq) := -(\mm M-\widehat{\mm M})\ddq-(\mm V_m-\widehat{\mm V_m})\dq -(\mm G-\widehat{\mm G}) -\mm F$ denotes the residual unknown term.

@@

The backstepping control-based approach is utilized to generate a reference signal $\mv z^*:= -{k_q}\tilde {\q}+\dq_d$ for $\mv {z}:= \dq$, where $\tilde {\q}:= \mv {q} - \qd$ and $k_q\in\R_{>0}$.
The desired stabilizing controller can be designed as follows:
\begin{equation}
    \mv \tau^* = 
    -\widehat{\mm M}\cdot ({k_z}\tilde {\mv z})
    + 
    ( 
        -\widehat{\mm M}\tilde {\q}
        +
        \widehat{\mm V_m}{\mv z}
        +
        \widehat{\mm G}
        -
        \mv f
        +
        \widehat{\mm M} \dot {\mv z}^*
    ) 
    ,
    \label{eq. desired control}
\end{equation}
where $\tilde {\mv z}:= \mv z-\mv z^*$ and $k_z\in\R_{>0}$.
Note that the desired controller cannot be realized because of $\mv f$.

To approximate the desired controller, an NN is utilized.
\color{red}
    Even though the DNNs have a higher approximation capability than the SHLNNs \cite{Rolnick:2018aa}, here, the SHLNN is utilized for simplicity and low computational complexity.
\color{black}
The NN with a single hidden layer is represented as 
\begin{equation}
    \NN({\q_{NN}};\wth):= \wV_1^\top \act(\wV_0^\top {\q_{NN}})
    ,
\end{equation}
where ${\q_{NN}}\in\R^{l_0+1}$ denotes the NN input vector, $\wV_i\in\R^{(l_i+1)\times l_{i+1}},\ i\in\{0,1\}$ denotes the weight matrix of the $i\textsuperscript{th}$ layer and $\act:\R^{l_1}\to\R^{l_1+1}$ denotes the activation function layer.
The activation function layer consists of an elementwise nonlinear function $\sigma(\cdot)$ and an augmented $1$ to combine the bias terms in the weight matrix (\ie $\act(\mv x) = \left(\sigma(\mv x_{(1)}),\cdots, \sigma(\mv x_{(l_1)}), 1\right)^\top $).
For further simplicity, let $\wth_i:= \myvec(\wV_i)\in\R^{\Xi_i}$ denotes the vectorized weights and $\wth:=(\wth_1^\top ,\wth_0^\top )^\top \in\R^{\Xi}$ denote the total weight vector, where $\Xi_i=(l_i+1)\cdot l_{i+1}$ and $\Xi=\Xi_0+\Xi_1$ denote the number of each layer and total weights, respectively.

Using this NN, the desired controller $\mv \tau^*$ can be approximated by the ideal weight vector $\wth^*$ for a compact subset $\Omega_{NN}\in\R^{l_0+1}$ to $\epsilon$ accuracy such that $\sup_{{\q_{NN}}\in\Omega_{NN}} \allowbreak \norm{\NN({\q_{NN}};\wth^*) - \mv \tau^*}=\epsilon<\infty$ \cite{Cybenko:1989aa}.
The ideal weight vector $\wth^*$ is typically assumed to be bounded.
Then, using the estimated weight vector $\estwth=(\estwth_1^\top ,\estwth_0^\top )^\top $ of $\wth^*=(\wth_1^{*\top},\wth_0^{*\top})^\top $, the desired controller $\mv \tau^*\approx -\NN(\q_{NN};\wth^*)-\mv\epsilon$ with bounded approximation error $\mv \epsilon\in\R^2$, can be approximated as follows:
\begin{equation}
    \mv \tau := -\NN({\q_{NN}};\estwth)
    .
    \label{eq. control law}
\end{equation}
For further sections, let $\idealNN:=\NN({\q_{NN}};\idealwth)$ and $\idealact :=\act({\idealwV}_0^{\top}{\q_{NN}})$, and let $\estNN:={\NN}({\q_{NN}};\estwth)$, $\estact :=\act(\estwV_0^{\top}{\q_{NN}})$ and $\estact' := \pptfrac{\estact}{(\estwV_0^\top \q_{NN})}$.

Using \eqref{eq. system dynamics 2}, \eqref{eq. desired control}, and \eqref{eq. control law}, the error dynamics can be obtained as the first-order system of augmented error $\mv \xi:=(\tilde {\q}^\top ,\tilde {\mv z}^\top)^\top \in\R^{2n}$ as follows:
\begin{equation}
    \dot{\mv \xi} 
    = 
    \mm A_\xi{\mv \xi}+\mm B_\xi(\NN^*-\estNN+\mv\epsilon)
    ,
\end{equation}
and 
\begin{equation}
    \mm A_\xi := 
    \begin{bmatrix}
        -{k_q} \mm I_{n} &\mm I_{n}
        \\
        -\mm I_{n}& -{k_z} \mm I_{n}
    \end{bmatrix}
    ,\ 
    \mm B_\xi := 
    \begin{bmatrix}
        \mm 0_{n\times n}\\\widehat{\mm M}^{-1}
    \end{bmatrix}.
\end{equation}

@@

\color{red}
    As discussed in Section \ref{sec:intro}, the boundedness of the weights should be considered to prevent excessively large control inputs and system instability.
\color{black}
For the boundedness of the weights, weight norm constraints are imposed on the adaptation process such that $c_{\theta_i}=\norm{{\estwth}_i}^2-\bar\theta^2_i,\ i\in\{0,1\}$.
The control problem can be reformulated into a constrained optimization problem as follows:
\begin{equation}
    \begin{matrix}
        \min_{{\estwth}} \ J(\mv \xi;{\estwth})
        :=
        \tfrac{1}{2}
        \mv \xi^\top \mm \Lambda \mv \xi
        \\ \\
        \begin{aligned}
        \text{subject to\;}&c_{j}\le0, \quad j\in\mathcal I 
        = 
        \{\theta_0,\theta_1\}
        ,
        \end{aligned}
    \end{matrix}
\end{equation}
where $\mm \Lambda=\mm \Lambda^\top >0$ denotes the weighting matrix.
In this optimization problem, $\mv \xi$ is considered a predefined parameter.
The corresponding Lagrangian function is defined as
\begin{equation}
    L(
        \mv \xi,
        {\estwth},
        [\lambda_j]_{j\in\mathcal I}
    )
    := 
    J(\mv \xi;{\estwth})
    +
    \textstyle\sum_{j\in\mathcal I} \lambda_jc_j({\estwth})
    .
\end{equation}
% where $\mathcal I=\{j\ \vert \ c_j\ge 0\}$ denotes the active set.
To solve the dual problem $\min_{{\estwth}} \max_{[\lambda_j]_{j\in\mathcal I}}  L$, the adaptation law is derived as follows:
\begin{subequations}
    \begin{align}
            \ddtt {{\estwth}}
            &
            =
            -\alpha 
            \pptfrac{L}{{\estwth}}
            =-\alpha 
            \left(
                \pptfrac{J}{{\estwth}}
                +
                \textstyle\sum_{j\in\mathcal{I}}
                \lambda_j 
                \pptfrac{c_j}{\estwth}
            \right),
        \label{eq. adaptation law th}
            \\
            \ddtt\lambda_j
            & 
            = 
            \beta_j
            \pptfrac{L}{\lambda_j} 
            = 
            \beta_j c_j ,
            \quad\quad\quad\quad      \      
            \forall j\in\mathcal I,
        \label{eq. adaptation law L}
            \\
            \lambda_j 
            & 
            = 
            \max(\lambda_j,0) ,
        \label{eq. adaptation law L max}
    \end{align}
    \label{eq. adaptation laws}
\end{subequations}
where arguments of $L$ are suppressed for brevity, and $\alpha\in\R_{>0}$ and $\beta_j\in\R_{>0}$ denote the adaptation gain and the update rate for each Lagrange multiplier, respectively.
\color{red}
% The Eq.~\eqref{eq. adaptation law L max} is analog to using an active set $\mathcal I=\{j\ \vert \ c_j\ge 0\}$ in constrained optimization theory which excludes terms of satisfied constraints in adaptation law.
\color{black}

Using the chain rule, the gradient of the objective function with respect to the weights (\ie $\pptfrac{J}{\estwth}$) in \eqref{eq. adaptation law th} can be represented as
\begin{equation}
    \pptfrac{J}{\estwth}
    =
    \pptfrac{\mv\xi}{\estwth}^\top 
    \mm \Lambda
    \mv\xi
    .
\end{equation}
The calculation of $\pptfrac{J}{\estwth}$ is not straightforward because of the dynamics of $\mv\xi$.
Using the forward sensitivity method presented in \cite{Sengupta:2014aa}, the sensitivity equation can be obtained as follows:
\begin{equation}
    \ddtt{\mm \eta} =
    \mm A_\xi \mm\eta
    - \mm B_\xi 
    % (\partial \estNN/\partial {\estwth})
    \pptfrac{\estNN}{\estwth}
    ,
    \label{eq. sen eq}
\end{equation}
where ${\mm \eta}:= \pptfrac{\mv\xi}{\estwth}\in\R^{2n\times \Xi}$ denotes the sensitivity of the weights to the augmented error.
The initial value of ${\mm \eta}$ is zero since the initial $\mv\xi$ is independent of the weights.
By decomposing for each layer, the dynamics of ${\mm \eta}_i:= \pptfrac{\mv\xi}{\estwth}_i\in\R^{2n\times \Xi_i}$ can be represented as
\begin{equation}
    \begin{aligned}
        \ddtt{\mm \eta} 
        \!=\!&
        \begin{bmatrix}
            {\mm \eta}_1&{\mm \eta}_0
        \end{bmatrix}'
        \\
        \!=\!&
        \mm A_\xi
        % \begin{bmatrix}
        [
            {\mm \eta}_1
            \ \ 
            {\mm \eta}_0
        ]
        % \end{bmatrix}
        \!-\!
        \mm B_\xi
        % \begin{bmatrix}
        [
            (\mm I_{l_2}\!\otimes\!\estact^\top )
            \ \ 
            \estwV_1^\top \estact'(\mm I_{l_1}\otimes {\q_{NN}}^\top )
        ]
        % \end{bmatrix}
        .
    \end{aligned}
\end{equation}
The calculation of $\pptfrac{\estNN}{\estwth}$ is introduced in \cite{Patil:2022aa}.
In conclusion, the gradient of the objective with respect to the weights can be obtained as $\pptfrac{J}{\estwth} = {\mm \eta}^\top\mm \Lambda\mv\xi$ by simulating the sensitivity equation \eqref{eq. sen eq}.

@@

The dynamics of $\mv \xi$ can be represented as
\begin{equation}    
    \ddtt {\mv\xi} = \mm A_\xi\mv \xi +\mm B_\xi
    (
        -\estwV_1^\top \estact+\mv w(t)
    )
    ,
\end{equation}
where $\mv w(t):= {\idealwV}_1^{\top}\idealact+\mv \epsilon$ is a lumped residual term, which is bounded as $\norm{\mv  w(t)}\le \bar w< 0$ since $\norm{\idealwth_1},\norm{\idealact}$ and $\norm{\mv\epsilon}$ are bounded.
The dynamics of ${\mm \eta}_1$ and ${\estwth}_1$ are represented as
\begin{equation}
    \begin{aligned}
        \ddtt{\mm \eta}_1 =
        & 
        \mm A_\xi {\mm \eta}_1 -\mm B_\xi (\mm I_{l_{2}}\otimes \estact^\top )
        ,
        \\
        \ddtt{{\estwth}}_1 =
        & -\alpha 
        (
            {\mm \eta}_1^\top 
            \mm \Lambda
            \mv\xi
            +
            2\lambda_{\theta_1} {\estwth}_1
        )
        .
    \end{aligned} 
\end{equation}
According to \cite[Chap.~4 T.~1.9]{Desoer:2009aa}, the boundedness of ${\mm \eta}_1$ can be obtained since $\mm A_\xi$ is stable and the residual term $-\mm B_\xi (\mm I_{l_{2}}\otimes \estact^\top )$ is bounded.

@@

The dynamics of ${\mm \eta}_0$ and ${\estwth}_0$ are represented as
\begin{equation}
    \begin{aligned}     
        \ddtt{\mm \eta}_0 =& 
        \mm A_\xi{\mm \eta}_0 -\mm B_\xi 
        \estwV_1^\top \estact'(\mm I_{l_1}\otimes {\q_{NN}}^\top )
        \\
        \ddtt{{\estwth}}_0
        =&
        -\alpha 
        (
            {\mm \eta}_0^\top \mm \Lambda\mv \xi+2\lambda_{\theta_0} {\estwth}_0
        )
        .
    \end{aligned}
\end{equation}
According to \cite[Chap.~4 T.~1.9]{Desoer:2009aa}, ${\mm \eta}_0$ is bounded since $\mm A_\xi$ is a stable matrix and $-\mm B_\xi\estwV_1^\top \estact'(\mm I_{l_1}\otimes {\q_{NN}}^\top )$ is bounded.
To obtain the invariance set of ${\estwth}_0$, taking the time derivative of the Lyapunov function $V_0=\tfrac{1}{2\alpha}{\estwth}_0^\top {\estwth}_0$ yields:
\begin{equation}
    \begin{aligned}
        \ddtt {V}_0 =& 
        -{\estwth}_0^\top (
            {\mm \eta}_0\mm \Lambda\mv \xi 
            +2\lambda_{\theta_0}{\estwth}_0
        )
        \\
        \le &
        \norm{{\estwth}_0} \norm{{\mm \eta}_0\mm \Lambda\mv \xi} 
        -
        2\lambda_{\theta_0} {\estwth}_0^\top {\estwth}_0
        ,
        \\
        \le &
        -2\lambda_{\theta_0} 
        \norm{{\estwth}_0}^2 
        + 
        \norm{{\mm \eta}_0\mm \Lambda\mv \xi}
        \norm{{\estwth}_0}
        .
    \end{aligned}
\end{equation}
Then, the invariance set can be represented as 
\begin{equation}
    \Theta_{{\theta}_0} :=
    \left\{
        {\estwth}_0\mid\norm{\estwth_0}
        \le
        \tfrac{
            \norm{{\mm \eta}_0\mm \Lambda\mv \xi}
        }{
            2\lambda_{\theta_0}
        }
    \right\}
    .    
\end{equation}
If $\lambda_{\theta_0}$ is increased sufficiently large due to the violation of $c_{\theta_0}$, the invariance set $\Theta_{{\theta}_0}$ converges to $\{{\estwth}_0 \mid \norm{{\estwth}_0} \le \bar\theta_0\}$ until the constraint $c_{\theta_0}$ is satisfied.
Therefore, the Lagrange multiplier $\lambda_{\theta_0}$ is also bounded.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\footnotesize
\bibliographystyle{ieeetr}
\bibliography{dding_template/refs}}

>>>>

\makeatletter
\let\NAT@parse\undefined
\makeatother

\def\pub{false} % true for publication, false for draft
\newcommand*{\template}{dding_template}
\input{\template/preamble/preamble_conf.tex}

@@

More recently, neuro–adaptive control approaches have been developed using neural networks (NNs) to approximate unknown system dynamics \cite{Farrell:2006aa}.
It generally uses the approximation capability of NNs with any arbitrary sigmoidal activation function \cite{Cybenko:1989aa}, which can approximate sufficiently smooth functions with minimum approximation error through ideal weights in a compact set.
This property allows the NNs to be synthesized in the controllers to approximate the unknown dynamics.
Various architectures of NNs have been proposed for neuro–adaptive control, such as single-hidden-layer NN (SHLNN) \cite{Esfandiari:2015aa,Gao:2006aa}, radial basis function NN (RBFNN) \cite{Ge:2002aa,Zhou:2023aa} and deep NN (DNN) \cite{Patil:2022aa}.
The literature has shown that NNs can improve the system's performance index and stability by approximating the unknown dynamics.

@@

% One of the common issues in neuro–adaptive control is that the boundedness of the control inputs is not guaranteed.
One of the common issues in using NNs 
\color{red}
for control compontnets
\color{black}
is that the outputs of NNs are not predictable.
This is because the input‒output relationships of the NNs are not interpretable (\ie the NNs are usually called black boxes \cite{Sheu:2020aa,Rudin:2019aa}).
\color{red}
This issue should be addressed for safety because the control input may reach the actuator's physical limitations and result in system collapse, due to the unpredictable control conponents from NNs.
\color{black}
\st{
    This issue should be addressed for safety because the unpredictable control input may reach the actuator's physical limitations and result in system collapse.
}

@@

Consider an uncertain Euler‒Lagrange system modeled as
\begin{equation}
    \mm M\ddq + \mm V_m\dq+ \mm G + \mm F 
    =
    \mv\tau 
    ,
    \label{eq. system dynamics 1}
\end{equation}
where $\q\in \mathbb{R}^n$ and $\mv \tau\in\R^n$ denote the generalized coordinate and the control input, respectively; $\mm M :=\mm M(\q)\in\mathbb{R}^{n\times n}$, $\mm V_m:= \mm V_m(\q,\dq)\in\mathbb{R}^{n\times n}$, and $\mm G:=\mm G(\q)\in\mathbb{R}^{n}$ denote the unknown system function matrices; and $\mm F:=\mm F(\dq)\in\mathbb{R}^{n}$ denotes the external force.
Using the user-designed matrices $\widehat{\mm M}>0,\widehat{\mm V}_m$ and $\widehat{\mm G}$, \eqref{eq. system dynamics 1} can be represented as 
\begin{equation}
    \widehat{\mm M}\ddq+\widehat{\mm V}_m\dq+\widehat{\mm G} = \mv \tau + \mv f(\q,\dq,\ddq)
    ,
    \label{eq. system dynamics 2}
\end{equation}
where $\mv f(\q,\dq,\ddq) := -(\mm M-\widehat{\mm M})\ddq-(\mm V_m-\widehat{\mm V}_m)\dq -(\mm G-\widehat{\mm G}) -\mm F$ denotes the residual unknown term.

@@

The backstepping control-based approach is utilized to generate a reference signal $\mv z^*:= -{k_q}\tilde {\q}+\dq_d$ for $\mv {z}:= \dq$, where $\tilde {\q}:= \mv {q} - \qd$ and $k_q\in\R_{>0}$.
The desired stabilizing controller can be designed as follows:
\begin{equation}
    \mv \tau^* = 
    -\widehat{\mm M}\cdot ({k_z}\tilde {\mv z})
    + 
    ( 
        -\widehat{\mm M}\tilde {\q}
        +
        \widehat{\mm V}_m{\mv z}
        +
        \widehat{\mm G}
        -
        \mv f
        +
        \widehat{\mm M} \dot {\mv z}^*
    ) 
    ,
    \label{eq. desired control}
\end{equation}
where $\tilde {\mv z}:= \mv z-\mv z^*$ and $k_z\in\R_{>0}$.
Note that the desired controller cannot be realized because of $\mv f$.

To approximate the desired controller, an NN is utilized.
\color{red}
    Even though the DNNs have a higher approximation capability than the SHLNNs \cite{Rolnick:2017aa}, here, the SHLNN is utilized for simplicity and low computational complexity.
\color{black}
The NN with a single hidden layer is represented as 
\begin{equation}
    \NN({\q_{n}};\wth):= \wV_1^\top \act(\wV_0^\top {\q_{n}})
    ,
\end{equation}
where ${\q_{n}}\in\R^{l_0+1}$ denotes the NN input vector, $\wV_i\in\R^{(l_i+1)\times l_{i+1}},\ \forall i\in\{0,1\}$ denotes the weight matrix of the $i\textsuperscript{th}$ layer and $\act:\R^{l_1}\to\R^{l_1+1}$ denotes the activation function layer.
The activation function layer consists of an elementwise nonlinear function $\sigma(\cdot)$ and an augmented $1$ to combine the bias terms in the weight matrix (\ie $\act(\mv x) = \left(\sigma(\mv x_{(1)}),\cdots, \sigma(\mv x_{(l_1)}), 1\right)^\top $).
For further simplicity, let $\wth_i:= \myvec(\wV_i)\in\R^{\Xi_i},\ \forall i\in\{0,1\}$ denotes the vectorized weights and $\wth:=(\wth_1^\top ,\wth_0^\top )^\top \in\R^{\Xi}$ denote the total weight vector, where $\Xi_i:=(l_i+1)\cdot l_{i+1},\ \forall i\in\{0,1\}$ and $\Xi:=\Xi_0+\Xi_1$ denote the number of each layer and total weights, respectively.

Using this NN, the desired controller $\mv \tau^*$ can be approximated by the ideal weight vector $\wth^*$ for a compact subset $\Omega_{NN}\in\R^{l_0+1}$ to $\epsilon$ accuracy such that $\sup_{{\q_{n}}\in\Omega_{NN}} \allowbreak \norm{\NN({\q_{n}};\wth^*) - \mv \tau^*}=\epsilon<\infty$ \cite{Cybenko:1989aa}.
The ideal weight vector $\wth^*$ is typically assumed to be bounded.
Then, using the estimated weight vector $\estwth:=(\estwth_1^\top ,\estwth_0^\top )^\top $ of $\idealwth:=({\idealwth_{1}}^{\top},{\idealwth_{0}}^{\top})^\top $ and bounded approximation error $\mv \epsilon\in\R^2$, the approximated desired controller $\mv \tau^*\approx -\NN(\q_{n};\idealwth)-\mv\epsilon$ can be estimated as follows:
\begin{equation}
    \mv \tau := -\NN({\q_{n}};\estwth)
    .
    \label{eq. control law}
\end{equation}
For further sections, let $\idealNN:=\NN({\q_{n}};\idealwth)$ and $\idealact :=\act({\idealwV_0}^{\top}{\q_{n}})$, and let $\estNN:={\NN}({\q_{n}};\estwth)$, $\estact :=\act(\estwV_0^{\top}{\q_{n}})$ and $\estact' := \pptfrac{\estact}{(\estwV_0^\top \q_{n})}$.

Using \eqref{eq. system dynamics 2}, \eqref{eq. desired control}, and \eqref{eq. control law}, the error dynamics can be obtained as the first-order system of augmented error $\mv \xi:=({\tilde {\q}}^\top ,{\tilde {\mv z}}^\top)^\top \in\R^{2n}$ as follows:
\begin{equation}
    \ddtt{\mv \xi} 
    = 
    \mm A_\xi{\mv \xi}+\mm B_\xi(\NN^*-\estNN+\mv\epsilon)
    ,
\end{equation}
and 
\begin{equation}
    \mm A_\xi := 
    \begin{bmatrix}
        -{k_q} \mm I_{n} &\mm I_{n}
        \\
        -\mm I_{n}& -{k_z} \mm I_{n}
    \end{bmatrix}
    ,\ 
    \mm B_\xi := 
    \begin{bmatrix}
        \mm 0_{n\times n}\\\widehat{\mm M}^{-1}
    \end{bmatrix}.
\end{equation}

@@

\color{red}
    As discussed in Section \ref{sec:intro}, the boundedness of the weights should be considered to prevent excessively large control inputs and system instability.
\color{black}
For the boundedness of the weights, weight norm constraints are imposed on the adaptation process such that $c_{\theta_i}:=c_{\theta_i}(\estwth_i)=\norm{{\estwth}_i}^2-\bar\theta^2_i,\ \forall i\in\{0,1\}$.
The control problem can be reformulated into a constrained optimization problem as follows:
\begin{equation}
    \begin{matrix}
        \min_{{\estwth}} \ J(\mv \xi;{\estwth})
        :=
        \tfrac{1}{2}
        \mv \xi^\top \mm \Lambda \mv \xi
        \\ \\
        \begin{aligned}
        \text{subject to\;}&c_{j}\le0, \quad j\in\mathcal I 
        := 
        \{\theta_0,\theta_1\}
        ,
        \end{aligned}
    \end{matrix}
\end{equation}
where $\mm \Lambda=\mm \Lambda^\top >0$ denotes the weighting matrix.
In this optimization problem, $\mv \xi$ is considered a predefined parameter.
The corresponding Lagrangian function is defined as
\begin{equation}
    L(
        \mv \xi,
        {\estwth},
        [\lambda_j]_{j\in\mathcal I}
    )
    := 
    J(\mv \xi;{\estwth})
    +
    \textstyle\sum_{j\in\mathcal I} \lambda_jc_j({\estwth})
    .
\end{equation}
% where $\mathcal I=\{j\ \vert \ c_j\ge 0\}$ denotes the active set.
To solve the dual problem $\min_{{\estwth}} \max_{[\lambda_j]_{j\in\mathcal I}}  L(\mv \xi,{\estwth},[\lambda_j]_{j\in\mathcal I}$, the adaptation law is derived as follows:
\begin{subequations}
    \begin{align}
            \ddtt {{\estwth}}
            &
            =
            -\alpha 
            \pptfrac{L}{{\estwth}}
            =-\alpha 
            \left(
                \pptfrac{J}{{\estwth}}
                +
                \textstyle\sum_{j\in\mathcal{I}}
                \lambda_j 
                \pptfrac{c_j}{\estwth}
            \right),
        \label{eq. adaptation law th}
            \\
            \ddtt\lambda_j
            & 
            = 
            \beta_j
            \pptfrac{L}{\lambda_j} 
            = 
            \beta_j c_j ,
            \quad\quad\quad\quad      \      
            \forall j\in\mathcal I,
        \label{eq. adaptation law L}
            \\
            \lambda_j 
            & 
            = 
            \max(\lambda_j,0) ,
        \label{eq. adaptation law L max}
    \end{align}
    \label{eq. adaptation laws}
\end{subequations}
where arguments of $L$ are suppressed for brevity, and $\alpha\in\R_{>0}$ and $\beta_j\in\R_{>0}$ denote the adaptation gain and the update rate for each Lagrange multiplier, respectively.
\color{red}
% The Eq.~\eqref{eq. adaptation law L max} is analog to using an active set $\mathcal I=\{j\ \vert \ c_j\ge 0\}$ in constrained optimization theory which excludes terms of satisfied constraints in adaptation law.
\color{black}

Using the chain rule, the gradient of the objective function with respect to the weights (\ie $\pptfrac{J}{\estwth}$) in \eqref{eq. adaptation law th} can be represented as
% \begin{equation}
$
    \pptfrac{J}{\estwth}
    =
    \pptfrac{\mv\xi}{\estwth}^\top 
    \mm \Lambda
    \mv\xi
    .
% \end{equation}
$
The calculation of $\pptfrac{J}{\estwth}$ is not straightforward because of the dynamics of $\mv\xi$.
Using the forward sensitivity method presented in \cite{Sengupta:2014aa}, the sensitivity equation can be obtained as follows:
\begin{equation}
    \ddtt{\mm \eta} =
    \mm A_\xi \mm\eta
    - \mm B_\xi 
    % (\partial \estNN/\partial {\estwth})
    \pptfrac{\estNN}{\estwth}
    ,
    \label{eq. sen eq}
\end{equation}
where ${\mm \eta}:= \pptfrac{\mv\xi}{\estwth}\in\R^{2n\times \Xi}$ denotes the sensitivity of the weights to the augmented error.
The initial value of ${\mm \eta}$ is zero since the initial $\mv\xi$ is independent of the weights.
By decomposing for each layer, the dynamics of ${\mm \eta}_i:= \pptfrac{\mv\xi}{\estwth}_i\in\R^{2n\times \Xi_i}$ can be represented as
\begin{equation}
    \begin{aligned}
        \ddtt{\mm \eta} 
        \!=\!&
        \begin{bmatrix}
            {\mm \eta}_1&{\mm \eta}_0
        \end{bmatrix}'
        \\
        \!=\!&
        \mm A_\xi
        % \begin{bmatrix}
        [
            {\mm \eta}_1
            \ \ 
            {\mm \eta}_0
        ]
        % \end{bmatrix}
        \!-\!
        \mm B_\xi
        % \begin{bmatrix}
        [
            (\mm I_{l_2}\!\otimes\!\estact^\top )
            \ \ 
            \estwV_1^\top \estact'(\mm I_{l_1}\otimes {\q_{n}}^\top )
        ]
        % \end{bmatrix}
        .
    \end{aligned}
\end{equation}
The calculation of $\pptfrac{\estNN}{\estwth}$ is introduced in \cite{Patil:2022aa}.
In conclusion, the gradient of the objective with respect to the weights can be obtained as $\pptfrac{J}{\estwth} = {\mm \eta}^\top\mm \Lambda\mv\xi$ by simulating the sensitivity equation \eqref{eq. sen eq}.

@@

The dynamics of $\mv \xi$ can be represented as
\begin{equation}    
    \ddtt {\mv\xi} = \mm A_\xi\mv \xi +\mm B_\xi
    (
        -\estwV_1^\top \estact+\mv w(t)
    )
    ,
\end{equation}
where $\mv w(t):= {\idealwV_1}^{\top}\idealact+\mv \epsilon$ is a lumped residual term, which is bounded as $\norm{\mv  w(t)}\le \bar w< 0$ since $\norm{\idealwth_1},\norm{\idealact}$ and $\norm{\mv\epsilon}$ are bounded.
The dynamics of ${\mm \eta}_1$ and ${\estwth}_1$ are represented as
\begin{equation}
    \begin{aligned}
        \ddtt{\mm \eta}_1 =
        & 
        \mm A_\xi {\mm \eta}_1 -\mm B_\xi (\mm I_{l_{2}}\otimes \estact^\top )
        ,
        \\
        \ddtt{{\estwth}}_1 =
        & -\alpha 
        (
            {\mm \eta}_1^\top 
            \mm \Lambda
            \mv\xi
            +
            2\lambda_{\theta_1} {\estwth}_1
        )
        .
    \end{aligned} 
\end{equation}
According to \cite[Chap.~4 T.~1.9]{Desoer:2009aa}, the boundedness of ${\mm \eta}_1$ can be obtained since $\mm A_\xi$ is stable and the residual term $\norm{-\mm B_\xi (\mm I_{l_{2}}\otimes \estact^\top )}$ is bounded.

@@

The dynamics of ${\mm \eta}_0$ and ${\estwth}_0$ are represented as
\begin{equation}
    \begin{aligned}     
        \ddtt{\mm \eta}_0 =& 
        \mm A_\xi{\mm \eta}_0 -\mm B_\xi 
        \estwV_1^\top \estact'(\mm I_{l_1}\otimes {\q_{n}}^\top )
        \\
        \ddtt{{\estwth}}_0
        =&
        -\alpha 
        (
            {\mm \eta}_0^\top \mm \Lambda\mv \xi+2\lambda_{\theta_0} {\estwth}_0
        )
        .
    \end{aligned}
\end{equation}
According to \cite[Chap.~4 T.~1.9]{Desoer:2009aa}, ${\mm \eta}_0$ is bounded since $\mm A_\xi$ is a stable matrix and $\norm{-\mm B_\xi\estwV_1^\top \estact'(\mm I_{l_1}\otimes {\q_{n}}^\top )}$ is bounded.
To obtain the invariance set of ${\estwth}_0$, taking the time derivative of the Lyapunov function $V_0=\tfrac{1}{2\alpha}{\estwth}_0^\top {\estwth}_0$ yields:
\begin{equation}
    \begin{aligned}
        \ddtt {V}_0 =& 
        -{\estwth}_0^\top (
            {\mm \eta}_0\mm \Lambda\mv \xi 
            +2\lambda_{\theta_0}{\estwth}_0
        )
        \\
        \le &
        \norm{{\estwth}_0} \norm{{\mm \eta}_0\mm \Lambda\mv \xi} 
        -
        2\lambda_{\theta_0} {\estwth}_0^\top {\estwth}_0
        ,
        \\
        \le &
        -2\lambda_{\theta_0} 
        \norm{{\estwth}_0}^2 
        + 
        \norm{{\mm \eta}_0\mm \Lambda\mv \xi}
        \norm{{\estwth}_0}
        .
    \end{aligned}
\end{equation}
Then, the invariance set can be represented as 
\begin{equation}
    \Theta_{{\theta}_0} :=
    \left\{
        {\estwth}_0\mid\norm{\estwth_0}
        \le
        \tfrac{
            \norm{{\mm \eta}_0\mm \Lambda\mv \xi}
        }{
            2\lambda_{\theta_0}
        }
    \right\}
    .    
\end{equation}
If $\lambda_{\theta_0}$ is increased sufficiently large due to the violation of $c_{\theta_0}$, the invariance set $\Theta_{{\theta}_0}$ converges to $\{{\estwth}_0 \mid \norm{{\estwth}_0} \le \bar\theta_0\}$ until the constraint $c_{\theta_0}$ is satisfied.
Therefore, the Lagrange multiplier $\lambda_{\theta_0}$ is also bounded.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\footnotesize
    \bibliographystyle{ieeetr}
    \bibliography{dding_template/refs}
}
